Case study: Genome-Wide Association Analyses (GWAS)

### 2018-07-05 datasets/GWAS/GWAS.Rmd
Add uninformative covariate to GWAS

# results files
resfile_N <- file.path(sbdir, paste0("bmi-samplesize-benchmark.rds"))
resfile_AF <- file.path(sbdir, paste0("bmi-maf-benchmark.rds"))

@@

@@

ggplot(data=bmi, aes(x = ind_covar_AF, y = SE)) +
  geom_hex(bins = 125)+
  xlab("Covariate: Minor allele frequency") 
```

@@

@@

@@

@@

```{r}
plotMethodRanks(c(resfile_AF, resfile_N), 
                colLabels = c("MAF", "Sample Size"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```

>>>>

# results files
resfile_N <- file.path(sbdir, paste0("bmi-samplesize-benchmark.rds"))
resfile_AF <- file.path(sbdir, paste0("bmi-maf-benchmark.rds"))
resfile_uninf <- file.path(sbdir, paste0("bmi-uninf-benchmark.rds"))

@@

Now we'll add an random (uninformative covariate) to compare results with.

```{r}
bmi <- bmi %>% mutate(ind_covar_uninf = rnorm(nrow(bmi)))
```

@@

ggplot(data=bmi, aes(x = ind_covar_AF, y = SE)) +
  geom_hex(bins = 125)+
  xlab("Covariate: Minor allele frequency")

ggplot(data=bmi, aes(x = ind_covar_AF, y = SE)) +
  geom_hex(bins = 125)+
  xlab("Covariate: Random") 
```

@@

### Covariate two: Random

Next we look at the random (uninformative) covariate.

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(bmi, pvalue="pval", covariate="ind_covar_uninf") +
  ggtitle("Covariate 1: Minor allele frequency")
```

```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(bmi, pvalue="pval", covariate="ind_covar_uninf", maxy=2)
```

@@

### Covariate three: Random

Now, we'll repeat the multiple testing correction using the other covariate (random):

```{r, GWAS1-sb3, results="hide", message=FALSE}
duration <- NA
if (!file.exists(resfile_uninf)) {
    t1 <- proc.time()
    sbU <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_uninf) %>%
                                    filter(!is.na(ind_covariate))), 
                              ftCols = c("ind_covar_uninf", "effect_size"),
                              parallel=TRUE, BPPARAM=multicoreParam)
    metadata(sbU)$data_download_link <-    
                     "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
    saveRDS(sbU, file = resfile_uninf)
    duration <- round((proc.time()-t1)[3]/60, 1)
} else {
    sbU <- readRDS(resfile_uninf)
}
```

@@

### Covariate three: Random

Next, we'll look at the performance metrics for the other covariate (random).

```{r, GWAS1-metrics3}
# rename assay to qvalue
assayNames(sbU) <- "qvalue"
sbU <- addDefaultMetrics(sbU)
```

```{r, GWAS1-plot3, results="hide"}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sbU, supplementary=FALSE) +
  ggtitle("Covariate 3: Random")

rejection_scatter_bins(sbU, covariate="ind_covar_uninf", bins=4,
                       supplementary=FALSE) +
  ggtitle("Covariate 3: Random")

# upset plot 
plotFDRMethodsOverlap(sbU, 
                      alpha=0.05, nsets=ncol(sbAF),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)
```

```{r, fig.width=8, fig.height=3.5}
covariateLinePlot(sbU, alpha=0.05, covname="effect_size", nbins=25, 
                 trans = "log1p")
covariateLinePlot(sbU, alpha=0.05, covname="ind_covar_uninf", nbins=25)
```

@@

```{r}
plotMethodRanks(c(resfile_AF, resfile_N, resfile_uninf), 
                colLabels = c("MAF", "Sample Size", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```
